---
output: 
  github_document:
    toc: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# The `rainfall` package

The `rainfall` package automatically downloads and processes rainfall data from 
the [Climatic Research Unit at the University of East Anglia](http://www.cru.uea.ac.uk/).

## Installation

The package is currently only available on GitHub. It can be installed through
the `devtools` package. 

```{r, echo = T, eval=FALSE}

devtools::install_github("felixhaass/rainfall")

```

## Functions & Output

The package uses basically only one function: `download_rain`. This functions
takes as input an ISO3 country code and downloads a file with monthly rainfall 
data for this country. 

It outputs a `data.frame` with the following variables. 

- `iso3c`: ISO3 code of a country
- `YEAR`: Year (currently 1901 - 2015)
- `month`: Month (character string of month, JAN - DEC)
- `month_numeric`: Month (numeric, 1-12)
- `rain`: Monthly mean of rainfall in milimeters per month (https://crudata.uea.ac.uk/cru/data/hrg/#info)
- `MAM, JJA, SON, DJF`: seasonal sum
- `ANN`: Annual sum of mean rainfall per month

For more information see `help(download_rain)`.

If you supply a vector of ISO3 country codes `download_rain()` will download
and combine the data for each country. 

## Usage

We download monthly rainfall data for the Democratic Republic of the Congo. To 
get the DRC's ISO3 code we use the excellent 
[`countrycode`](https://cran.r-project.org/web/packages/countrycode/index.html) 
package.


```{r, echo = T, eval=T, warning = F, message=F}

library(countrycode)
library(rainfall)

country_iso3_list <- countrycode("Democratic Repulic of the Congo", 
                                 "country.name",
                                 "iso3c")

drc_rain <- download_rain(country = country_iso3_list, 
                          long = T, delete_raw = T)


```


This produces the following `data.frame` named `drc_rain` 
(only last 12 rows are shown):


```{r, echo = F, eval=T}

knitr::kable(tail(drc_rain, 12))

```

The same works for multiple countries:


```{r, echo = T, eval=T, warning = F, message=F}


# Note that we add several countries to the list
country_list <- c("Democratic Repulic of the Congo",
                  "Germany", 
                  "Philippines", 
                  "Algeria")

country_iso3_list <- countrycode(country_list, 
                                 "country.name",
                                 "iso3c")

compare_rain <- download_rain(country = country_iso3_list, 
                          long = T, delete_raw = T)


```

We can now plot comparison of rainfall in these countries in 2015:


```{r rain_comparison, dpi = 200, echo = T, eval=T, warning = F, message=F, fig.path = "./example_images/"}

library(ggplot2)
library(dplyr)

compare_rain_plot <- compare_rain %>% 
  filter(YEAR == 2015) %>% 
  ggplot(., 
         aes(x = month_numeric, y = rain, group =iso3c)) +
  geom_line(aes(color = iso3c)) +
  scale_color_brewer(palette = "RdYlGn") +
  theme_bw() +
  scale_x_continuous(breaks=1:12, labels = month.abb) +
  labs(x = "", y = "Monthly Average Rainfall")

print(compare_rain_plot)


```

## More information on the data

I did *not* create this data! For more information on the data, see the original
research article describing the data:

> Harris, I, PD Jones, TJ Osborn & DH Lister (2014) [Updated high-resolution grids of monthly climatic observations - the CRU TS3.10 Dataset.](http://onlinelibrary.wiley.com/doi/10.1002/joc.3711/abstract) International Journal of Climatology 34(3): 623-642.

## Feedback & Questions
For feedback, questions, and problems regarding the package, please use GitHub
[Issues.](https://github.com/felixhaass/rainfall/issues).
